---
title: "Estimate Characteristics"
author: "Ellie Mainou"
date: "4/6/2021"
output: html_document
---

Estimate 
2. Peak VL magnitude
3. Peak VL time
4. Peak VL joint measure (i.e. magnitude+time)


for all four models: 1. Standard 2. DDDI 3. Macro 4. MacroDDDI

For growth rate: run the model to equilibrium. The equilibrium is the model-derived setpoint. Peak VL magnitude is just the max viral load measurement of that model, Peak VL time is the time post-positive test when the peak magnitude occurs. Peak VL joint is a measure that tries to capture the magnitude and time of the peak jointly. 


```{r}
library(readxl)
library(readr)
library(deSolve)
```

```{r}
#Import modified dataset and make sure that it reads as numbers
NewAcuteData <- read_csv("HIV R files/Acute Infection Models/RV217ForModels5.csv")
#Matrices with parameter values and AIC, BIC, AICc
StandardFittedPrms <- read_csv("HIV R files/Acute Infection Models/Model Fitting/StandardFittedPrms_5.csv")
DDDIFittedPrms <- read_csv("HIV R files/Acute Infection Models/Model Fitting/DDDIFittedPrms_4.csv")
MacroFittedPrms <- read_csv("HIV R files/Acute Infection Models/Model Fitting/MacroFittedPrms_4.csv")
MacroDDDIFittedPrms <- read_csv("HIV R files/Acute Infection Models/Model Fitting/MacroDDDIFittedPrms_3.csv")
PrmErrorMatrixRV217_6 <- read_csv("HIV R files/RV217 Data Analysis /PrmErrorMatrixRV217_6.csv")
Peak_RV217_6 <- read_csv("HIV R files/RV217 Data Analysis /Peak_RV217_6.csv")
```



```{r}
ID<-unique(NewAcuteData$ID)  #participant IDs
observations= length(ID)
```


```{r}
# Model Equations
Hstart=1e6
b=0.01
d=0.01
T.start=1e6

Model.Standard <- function(t, y, parms) { 
  with(as.list(c(y,parms)),{  
    k <- ( (delta+r)*(c+r) ) / ( 10^log10p*T.start )
    dT <- d*T.start -d*T - k*T*V 
    dI <- k*T*V - delta*I
    dV <- 10^log10p*I - c*V 
    dy=c(dT, dI, dV)
    return(list(dy))
  }) 
} 

Model.DDDI <- function(t, y, parms) { 
  with(as.list(c(y,parms)),{  
    k <- ( r*(c+r) ) / ( 10^log10p*T.start )
    dT <- d*T.start -d*T - k*T*V 
    dI <- k*T*V - delta*I^(gamma+1)
    dV <- 10^log10p*I - c*V 
    dy=c(dT, dI, dV)
    return(list(dy))
  }) 
}

Model.Macro <- function(t, y, parms) { 
  with(as.list(c(y,parms)),{ 
    beta=(r+eta)*(r+alpha+b)/Hstart/(10^log10lambda-r-alpha-b)
    dH <- b*Hstart - b*H - alpha*P 
    dV <- 10^log10lambda*P - eta*V - beta*H*V
    dP <-  beta*H*V - b*P - alpha*P - alpha*(1+k)/k*P^2/H 
    dy=c(dH,dV,dP)
    return(list(dy))
  }) 
} 

Model.MacroDDDI <- function(t, y, parms) { 
  with(as.list(c(y,parms)),{ 
  if ((gamma+1)>1){  #it should be 1 but in the equation I use gamma+1 so I can make the lower bound zero
    beta=(r+eta)*(r+b)/Hstart/(10^log10lambda-r-b)
  }else if((gamma+1)==1){ beta=(r+eta)*(r+b+alpha)/Hstart/(10^log10lambda-r-b-alpha)
  }    
    dH <- b*Hstart - b*H -  alpha^(gamma+1)*(P^(gamma+1))/H^(gamma+1-1)
    dV <- 10^log10lambda*P - eta*V - beta*H*V
    dP <-  beta*H*V - b*P - (alpha^(gamma+1))*((P^(gamma+1))/H^(gamma+1-1))*(1+ (P/H)*(1+k)/k)^(gamma+1)
    dy=c(dH,dV,dP)
    return(list(dy))
  }) 
}
```


```{r}
#Beta for Macro and MacroDDDI

MacroFittedPrms$beta=(MacroFittedPrms$r+MacroFittedPrms$eta)*(MacroFittedPrms$r+MacroFittedPrms$alpha+b)/Hstart/(10^MacroFittedPrms$log10lambda-MacroFittedPrms$r-MacroFittedPrms$alpha-b)


MacroDDDIFittedPrms$beta=(MacroDDDIFittedPrms$r+MacroDDDIFittedPrms$eta)*(MacroDDDIFittedPrms$r+b+MacroDDDIFittedPrms$alpha)/Hstart/(10^MacroDDDIFittedPrms$log10lambda-MacroDDDIFittedPrms$r-b-MacroDDDIFittedPrms$alpha)

```


```{r}
# Matrices with growth, peak, decay, setpoint and squared errors among models 
names<-c("ID", "PeakM", "Standard", "SQ_Standard", "DDDI", "SQ_DDDI", "Macro", "SQ_Macro", "MacroDDDI", "SQ_MacroDDDI",  "Model")
ComparePeakM<-as.data.frame(matrix(data=NA, nrow=observations, ncol=length(names)))
colnames(ComparePeakM)<-names
ComparePeakM$ID<-ID

names<-c("ID", "PeakT", "Standard", "SQ_Standard", "DDDI", "SQ_DDDI", "Macro", "SQ_Macro", "MacroDDDI", "SQ_MacroDDDI",  "Model")
ComparePeakT<-as.data.frame(matrix(data=NA, nrow=observations, ncol=length(names)))
colnames(ComparePeakT)<-names
ComparePeakT$ID<-ID

names<-c("ID", "PeakJ", "Standard", "SQ_Standard", "DDDI", "SQ_DDDI", "Macro", "SQ_Macro", "MacroDDDI", "SQ_MacroDDDI", "Model")
ComparePeakJ<-as.data.frame(matrix(data=NA, nrow=observations, ncol=length(names)))
colnames(ComparePeakJ)<-names
ComparePeakJ$ID<-ID


```


```{r}
#Round t0 to the first decimal place
StandardFittedPrms$t0<-round(StandardFittedPrms$t0, 1)
DDDIFittedPrms$t0<-round(DDDIFittedPrms$t0, 1)
MacroFittedPrms$t0<-round(MacroFittedPrms$t0, 1)
MacroDDDIFittedPrms$t0<-round(MacroDDDIFittedPrms$t0, 1)
```

```{r}
for (i in 1:observations){
  #Data
  patient.current<-NewAcuteData[which(NewAcuteData$ID==ID[i]), ]
  time.points<-patient.current$days
  log10vl<-patient.current$log10VL
  vldata <- data.frame(time.points,log10vl)
  n=nrow(vldata)
  
#Run ODES
  
  #Standard
  init.cond=c(T=T.start, I=StandardFittedPrms$c[i]*1/(10^StandardFittedPrms$log10p[i]), V=1)


  parameters.Staff=list(
  d=d,
  T.start=T.start,
  log10p=StandardFittedPrms$log10p[i],
  delta=StandardFittedPrms$delta[i],
  c=StandardFittedPrms$c[i],
  r=StandardFittedPrms$r[i],
  t0=StandardFittedPrms$t0[i]
  )

soln.Standard=as.data.frame(ode(y=init.cond, times=c(seq(StandardFittedPrms$t0[i], 0, 0.01), 0, seq(0.01, 100, 0.01)), Model.Standard, parms=parameters.Staff, hmax=0.01))

#DDDI
init.cond=c(T=T.start, I=DDDIFittedPrms$c[i]*1/(10^DDDIFittedPrms$log10p[i]), V=1)


  parameters.Hol=list(
  d=d,
  T.start=T.start,
  log10p=DDDIFittedPrms$log10p[i],
  delta=DDDIFittedPrms$delta[i],
  c=DDDIFittedPrms$c[i],
  r=DDDIFittedPrms$r[i],
  gamma=DDDIFittedPrms$gamma[i],
  t0=DDDIFittedPrms$t0[i]
  )

  soln.DDDI=as.data.frame(ode(y=init.cond, times=c(seq(DDDIFittedPrms$t0[i], 0, 0.01), 0, seq(0.01, 100, 0.01)), Model.DDDI, parms=parameters.Hol, hmax=0.01))


#Macro
init.cond=c(H=Hstart, V=1, P=0)

  parameters.Koel=list(
  b=b,
  Hstart=Hstart,
  log10lambda=MacroFittedPrms$log10lambda[i],
  alpha=MacroFittedPrms$alpha[i],
  eta=MacroFittedPrms$eta[i],
  k=MacroFittedPrms$k[i],
  r=MacroFittedPrms$r[i],
  t0=MacroFittedPrms$t0[i]
  )

  soln.Macro=as.data.frame(ode(y=init.cond, times=c(seq(MacroFittedPrms$t0[i], 0, 0.01), seq(0.01, 100, 0.01)), Model.Macro, parms=parameters.Koel, hmax=0.01))

    #MacroDDDI
  init.cond=c(H=Hstart, V=1, P=0)

  parameters.KoelD=list(
  b=b,
  Hstart=Hstart,
  log10lambda=MacroDDDIFittedPrms$log10lambda[i],
  alpha=MacroDDDIFittedPrms$alpha[i],
  eta=MacroDDDIFittedPrms$eta[i],
  k=MacroDDDIFittedPrms$k[i],
  r=MacroDDDIFittedPrms$r[i],
  gamma=MacroDDDIFittedPrms$gamma[i],
  t0=MacroDDDIFittedPrms$t0[i]
  )

  soln.MacroDDDI=as.data.frame(ode(y=init.cond, times=c(seq(MacroDDDIFittedPrms$t0[i], 0, 0.01), seq(0.01, 100, 0.01)), Model.MacroDDDI, parms=parameters.KoelD, hmax=0.01))
  
ComparePeakM$Standard[i]<-max(log10(soln.Standard$V))
ComparePeakM$DDDI[i]<-max(log10(soln.DDDI$V))
ComparePeakM$Macro[i]<-max(log10(soln.Macro$V))
ComparePeakM$MacroDDDI[i]<-max(log10(soln.MacroDDDI$V))

ComparePeakT$Standard[i]=round(soln.Standard$time[which(soln.Standard$V==max(soln.Standard$V))], 2)
ComparePeakT$DDDI[i]=round(soln.DDDI$time[which(soln.DDDI$V==max(soln.DDDI$V))], 2)
ComparePeakT$Macro[i]=round(soln.Macro$time[which(soln.Macro$V==max(soln.Macro$V))], 2)
ComparePeakT$MacroDDDI[i]=round(soln.MacroDDDI$time[which(soln.MacroDDDI$V==max(soln.MacroDDDI$V))], 2)

ComparePeakJ$Standard[i]= sqrt((ComparePeakM$Standard[i])^2+ (ComparePeakT$Standard[i])^2)
ComparePeakJ$DDDI[i]= sqrt((ComparePeakM$DDDI[i])^2+ (ComparePeakT$DDDI[i])^2)
ComparePeakJ$Macro[i]= sqrt((ComparePeakM$Macro[i])^2+ (ComparePeakT$Macro[i])^2)
ComparePeakJ$MacroDDDI[i]= sqrt((ComparePeakM$MacroDDDI[i])^2+ (ComparePeakT$MacroDDDI[i])^2)


if(!is.na(Peak_RV217_6$actual_peak_vl[which(Peak_RV217_6$ID==ID[i])])){
#Peak Magnitude
ComparePeakM$PeakM[i]<-Peak_RV217_6$actual_peak_vl[which(Peak_RV217_6$ID==ID[i])]


#Find which model predicts peak viral load magnitude better
x0<-(ComparePeakM$PeakM[i]-ComparePeakM$Standard[i])^2
ComparePeakM$SQ_Standard[i]<-x0
x1<-(ComparePeakM$PeakM[i]-ComparePeakM$DDDI[i])^2
ComparePeakM$SQ_DDDI[i]<-x1
x2<-(ComparePeakM$PeakM[i]-ComparePeakM$Macro[i])^2
ComparePeakM$SQ_Macro[i]<-x2
x3<-(ComparePeakM$PeakM[i]-ComparePeakM$MacroDDDI[i])^2
ComparePeakM$SQ_MacroDDDI[i]<-x3


if( min(x0, x1, x2,x3)==x0){
  ComparePeakM$Model[i]<-"Standard"
  }else if ( min(x0, x1, x2,x3)==x1){
      ComparePeakM$Model[i]<-"DDDI"
    }else if ( min(x0, x1, x2,x3)==x2){
      ComparePeakM$Model[i]<-"Macro"
    }else if ( min(x0, x1, x2,x3)==x3){
      ComparePeakM$Model[i]<-"MacroDDDI"
    }

#Peak Time
ComparePeakT$PeakT[i]=Peak_RV217_6$actual_peak_time[which(Peak_RV217_6$ID==ID[i])]

x0<-(ComparePeakT$PeakT[i]-ComparePeakT$Standard[i])^2
ComparePeakT$SQ_Standard[i]<-x0
x1<-(ComparePeakT$PeakT[i]-ComparePeakT$DDDI[i])^2
ComparePeakT$SQ_DDDI[i]<-x1
x2<-(ComparePeakT$PeakT[i]-ComparePeakT$Macro[i])^2
ComparePeakT$SQ_Macro[i]<-x2
x3<-(ComparePeakT$PeakT[i]-ComparePeakT$MacroDDDI[i])^2
ComparePeakT$SQ_MacroDDDI[i]<-x3


#Find which model predicts peak viral load time better
if( min(x0, x1, x2,x3)==x0){
  ComparePeakT$Model[i]<-"Standard"
  }else if ( min(x0, x1, x2,x3)==x1){
      ComparePeakT$Model[i]<-"DDDI"
    }else if ( min(x0, x1, x2,x3)==x2){
      ComparePeakT$Model[i]<-"Macro"
    }else if ( min(x0, x1, x2,x3)==x3){
      ComparePeakT$Model[i]<-"MacroDDDI"
    }

#Peak Joint, i.e. both the magnitude and the time; for that use a euclidean distance from the origin. Need to account for different magnitudes between peak and time--(time/2)
#from data
ComparePeakJ$PeakJ[i]=Peak_RV217_6$actual_peak_vl[which(Peak_RV217_6$ID==ID[i])]+(Peak_RV217_6$actual_peak_time[which(Peak_RV217_6$ID==ID[i])])


x0<-(ComparePeakJ$PeakJ[i]-ComparePeakJ$Standard[i])^2
ComparePeakJ$SQ_Standard[i]<-x0
x1<-(ComparePeakJ$PeakJ[i]-ComparePeakJ$DDDI[i])^2
ComparePeakJ$SQ_DDDI[i]<-x1
x2<-(ComparePeakJ$PeakJ[i]-ComparePeakJ$Macro[i])^2
ComparePeakJ$SQ_Macro[i]<-x2
x3<-(ComparePeakJ$PeakJ[i]-ComparePeakJ$MacroDDDI[i])^2
ComparePeakJ$SQ_MacroDDDI[i]<-x3

#Find which model predicts peak  joint better
if( min(x0, x1, x2,x3)==x0){
  ComparePeakJ$Model[i]<-"Standard"
  }else if ( min(x0, x1, x2,x3)==x1){
      ComparePeakJ$Model[i]<-"DDDI"
    }else if ( min(x0, x1, x2,x3)==x2){
      ComparePeakJ$Model[i]<-"Macro"
    }else if ( min(x0, x1, x2,x3)==x3){
      ComparePeakJ$Model[i]<-"MacroDDDI"
    }
}

}

write.csv(ComparePeakM, file="ComparePeakM.csv")
write.csv(ComparePeakT, file="ComparePeakT.csv")
write.csv(ComparePeakJ, file="ComparePeakJ.csv")
```
